---
title: "Research Overview"
author: "Susan Vanderplas"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    includes:
      after_body:
        "js-addins.html"
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["default", "default-fonts", "css/this-presentation.css", "css/modal.css"]
    lib_dir: libs
    nature:
      countIncrementalSlides: false
---

```{r, include = F, eval = T}
knitr::opts_chunk$set(echo=F, dpi=300)
library(tidyverse)
clean_file_name <- function(x) {
  basename(x) %>% str_remove("\\..*?$") %>% str_remove_all("[^[A-z0-9_]]")
}

img_modal <- function(src, alt = "", id = clean_file_name(src), other = "") {
  
  other_arg <- paste0("'", as.character(other), "'") %>%
    paste(names(other), ., sep = "=") %>%
    paste(collapse = " ")
  
  js <- glue::glue("<script>
        /* Get the modal*/
          var modal{id} = document.getElementById('modal{id}');
        /* Get the image and insert it inside the modal - use its 'alt' text as a caption*/
          var img{id} = document.getElementById('img{id}');
          var modalImg{id} = document.getElementById('imgmodal{id}');
          var captionText{id} = document.getElementById('caption{id}');
          img{id}.onclick = function(){{
            modal{id}.style.display = 'block';
            modalImg{id}.src = this.src;
            captionText{id}.innerHTML = this.alt;
          }}
          /* When the user clicks on the modalImg, close it*/
          modalImg{id}.onclick = function() {{
            modal{id}.style.display = 'none';
          }}
</script>")
  
  html <- glue::glue(
     " <!-- Trigger the Modal -->
<img id='img{id}' src='{src}' alt='{alt}' {other_arg}>

<!-- The Modal -->
<div id='modal{id}' class='modal'>

  <!-- Modal Content (The Image) -->
  <img class='modal-content' id='imgmodal{id}'>

  <!-- Modal Caption (Image Text) -->
  <div id='caption{id}' class='modal-caption'></div>
</div>
"
  )
  write(js, file = "js-addins.html", append = T)
  return(html)
}

# Clean the file out at the start of the compilation
write("", file = "js-addins.html")
```


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

## Graphics and Visualization

- Experiments in graphical perception

```{r results='asis', echo = F, include = T}
cat(img_modal(src = "sine-illusion-app.png", alt = "Graphical Optical Illusions"))
```

---
## Graphics and Visualization

.pull-left[
- Experiments in graphical perception

- Using graphics for testing statistical hypotheses

].pull-right[
```{r results='asis', echo = F, include = T}
cat(img_modal(src = "feature-hierarchy-color-ellipse-11-20.png", alt = "Testing hypotheses using graphs"))
```
]

---
## Graphics and Visualization

.pull-left[
- Experiments in graphical perception

- Using graphics for testing statistical hypotheses

].pull-right[
```{r results='asis', echo = F, include = T}
cat(img_modal(src = "feature-hierarchy-line-11-20.png", alt = "Testing hypotheses using graphs"))
```
]

---
## Graphics and Visualization

.pull-left[
- Experiments in graphical perception

- Using graphics for testing statistical hypotheses

- Historical graphics -     
exploration, re-creation with modern methods, testing
].pull-right[

```{r results='asis', echo = F, include = T}
cat(img_modal(src = "statistical-atlas-church-accommodation.png", alt = "Historical Graphics - 1874 Statistifcal Atlas"))
cat(img_modal(src = "Michigan-spine_with_frame1.png", alt = "Recreating plots from the 1874 Statistifcal Atlas", other=list(width="49%")))
cat(img_modal(src = "Michigan-spine_without_frame1.png", alt = "Recreating plots from the 1874 Statistifcal Atlas", other=list(width="49%")))
```
]
---

## Graphics and Visualization

Designing graphics for unique data sets: large, complicated, messy...


```{r results='asis', echo = F, include = T}
cat(c(img_modal(src = "SNPAggregateBrowser.png", alt = "Soybean Line Genomic Study: Aggregate SNP Distribution", other = list(width = "30%")),
img_modal(src = "SNPInheritance.png", alt = "Soybean Line Genomic Study: SNP Inheritance", other = list(width = "30%")),
img_modal(src = "SNPKinship.png", alt = "Soybean Line Genomic Study: SNP Kinship and Phenotype", other = list(width = "30%"))))
```

---

## Forensics

Statistical Matching of Bullets

.pull-left[
```{r results='asis', echo = F, include = T}
cat(img_modal(src = "HS36-Bullet-With-Crosscut.png", alt = "Bullet land scan, with crosscut"))
cat(img_modal(src = "Combined-cross-section.png", alt = "Crosscut and profile"))
```
].pull-right[
```{r results='asis', echo = F, include = T}
cat(img_modal(src = "signature-combined.png", alt = "Bullet smooth and resulting signature"))
cat(img_modal(src = "signature-aligned.png", alt = "Aligned bullet signatures"))
```

]

---

## Forensics

Identification of Shoe Features

```{r results='asis', echo = F, include = T}
cat(img_modal(src = "similar-shoes.png", alt = "Geometric features on shoes", other=list(width="80%")))
```