---
title: "Shoeprint Identification"
author: "Susan VanderPlas"
output:
  rmdshower::shower_presentation:
    self_contained: false
    ratio: 16x10
    css: shower.css
---

```{r setup, cache = F, echo = F, include = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T)

library(tidyverse)
library(ShoeprintCleanR)
library(gridExtra)
library(imager)

pic_paths <- list.files("/storage/CSAFE/Test Shoes/Prints/", 
                        full.names = T, 
                        pattern = "(\\d{1,}[LR]_\\d{8}_\\d_\\d_\\d_csafe_[a-z]{1,}).tif$")
i <- which(str_detect(pic_paths, "006070R_20171113_6_1_1_csafe_jekruse"))
orig_shoe <- shoe <- load.image(pic_paths[i])
```
# Matching Footwear Impressions

## (Ideal version) Original Data
```{r out.width = "90%", fig.width = 6, fig.height = 4}
cleanshoe <- shoe %>%
  quantize_colors(8) %>%
  remove_print_label(thr = "10%") %>%
  remove_local_background() %>%
  quantize_colors(4) %>%
  remove_border_lines(maxiter = 10) %>%
  crop_border(axis = "xy", tol = .1) %>%
  align_shoe_print() %>%
  crop_border()

plot(imrotate(shoe, 90), axes = F)
# plot(imrotate(cleanshoe, 90), axes = F)
```


## (Ideal version) Cleaned Data
```{r out.width = "90%", fig.width = 6, fig.height = 4}
# plot(imrotate(shoe, 90), axes = F)
plot(imrotate(cleanshoe, 90), axes = F)
```
Intelligent automatic processing of images to remove background artifacts

## (Ideal version) Analysis

- Identify important features of the shoe
- Extract the features
- Compare to a database of shoeprint features or a set of suspect shoe prints
- Find a candidate set of shoes to examine further
- Use image matching algorithms to identify correspondance between test image and database images
- Determine probability of match
- Compare to current methods (human perception)


## Less Ideal version

- Partial prints
- Smears
- Incomplete databases
- Hard to identify shoe model from print or partial print to get comparison prints

## Computational Methods

- Random Forests
- Image analysis methods- SIFT, SURF, etc.
- ???
- R package development to make code available to crime labs and practitioners

## Data size

- 15,315 images (color, black and white, smudged, unsmudged, 2D, 3D) 
    - growing weekly
- 4902 black and white powder prints 
    - = 207 GB